\documentclass{article}

\usepackage{courier}
\usepackage{graphicx} % Required for the inclusion of images
\usepackage{listings}
\usepackage{color}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{float}
\usepackage[toc,page]{appendix}
\usepackage{dcolumn}
\usepackage{pdflscape}
\usepackage{hyperref}
\usepackage{framed}
\usepackage[english]{babel}
\usepackage{soul}
\usepackage{caption}
\usepackage{amsfonts}
\usepackage{MnSymbol,wasysym}
\usepackage{algorithmic}
\usepackage{mathtools}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\captionsetup[figure]{labelformat=empty}%

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\renewcommand{\labelenumi}{\alph{enumi}.} % Make numbering in the enumerate environment by letter rather than number (e.g. section 6)

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=R,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\newcommand{\footlabel}[2]{%
    \addtocounter{footnote}{1}%
    \footnotetext[\thefootnote]{%
        \addtocounter{footnote}{-1}%
        \refstepcounter{footnote}\label{#1}%
        #2%
    }%
    $^{\ref{#1}}$%
}

\newcommand{\footref}[1]{%
    $^{\ref{#1}}$%
}

%\usepackage[colorlinks]{hyperref}
\hypersetup{linkcolor=DarkRed}
\hypersetup{urlcolor=DarkBlue}
\usepackage{cleveref}

\title{Data Mining\\Homework Assignment \#12} % Title

\author{Dmytro Fishman, Anna Leontjeva and Jaak Vilo} % Author name

\begin{document}

\maketitle % Insert the title, author and date 
The goal for this homework is to get you started with basics of social network analysis. 

\section*{Task 1}
State why for the integration of multiple heterogeneous information sources, many
companies in industry prefer the update-driven approach (which constructs and uses
data warehouses), rather than the query-driven approach (which applies wrappers and
integrators). Describe situations where the query-driven approach is preferable over
the update-driven approach.

\section*{Task 2}
A data warehouse can be modeled by either a star schema or a snowﬂake schema.
Brieﬂy describe the similarities and the differences of the two models, and then
analyze their advantages and disadvantages with regard to one another. Give your
opinion of which might be more empirically useful and state the reasons behind
your answer.

\section*{Task 3}
A data warehouse can be modeled by either a star schema or a snowﬂake schema.
Brieﬂy describe the similarities and the differences of the two models, and then
analyze their advantages and disadvantages with regard to one another. Give your
opinion of which might be more empirically useful and state the reasons behind
your answer.
 
\section*{Task 4}
Suppose that a data warehouse consists of the four dimensions, date, spectator, location, and game, and the two measures, count and charge, where charge is the fare that
a spectator pays when watching a game on a given date. Spectators may be students,
adults, or seniors, with each category having its own charge rate.
(a) Draw a star schema diagram for the data warehouse.
(b) Starting with the base cuboid [date, spectator, location, game], what speciﬁc OLAP
operations should one perform in order to list the total charge paid by student
spectators at GM Place in 2004?

\section*{Task 5}
Explain the terms slice and dice, drill-down and roll-up. Provide examples, e.g. from the census data analysis.

\section*{Task 6}
In this bonus task you will help biologists to fight the virus from the task one and two. Use data email\_virus.txt, calculate the same components as in task 2b and the probability of emerging a 'large scale' epidemics as in task 2c. Hint: you may want to use BFS algorithm to estimate IN and OUT components.
\medskip
\\
This homework was inspired by Juri Leskovec stanford course ''Social and Information Network Analysis``. 
\end{document}
